---
title: "Exercise sheet 1: Nussinov algorithm"
---

---------------------------------

# Exercise 1

Consider the RNA with sequence:

```
S = GGGCACAUGGGGCAGUGCAGCCACUGAGCC
```

$P = \{(1,30),(2,29),(4,17),(5,16),(6,15),(8,14),(9,13),(18,26),(19,25),(20,24)\}$


### 1.

::: {.question data-latex=""}

Draw the structure in dot-bracket notation and as a graphical representation.

:::

#### {.tabset}

##### Hide


##### Solution

::: {.answer data-latex=""}

```{r, include=knitr::is_html_output(), echo=FALSE,}
knitr::include_graphics("assets/figures/exercise-sheet-1/graph_represention_part1.svg")
```

:::

#### {-}

---------------------------------

### 2.

::: {.question data-latex=""}

Consider the structure $P'=P\cup\{(11,22),(12,21)\}$ of sequence S.

Modify your drawings, where possible, in order to show the new structure $P'$. Where does
this fail and why?

:::

#### {.tabset}

##### Hide

##### Solution

::: {.answer data-latex=""}
Problem: pseudoknot

```{r, include=knitr::is_html_output(), echo=FALSE,}
knitr::include_graphics("assets/figures/exercise-sheet-1/graph_represention_part2.svg")
```

The new base pairs/arcs $(11, 22)$ and $(12, 21)$ result in a pseudoknot.
The bracket-only dot-bracket notation can not display pseudoknots.
One would need to introduce a different symbol eg using square brackets '\texttt{[]}' to indicate the opening and closing positions of the crossing base pairs.
:::

#### {-}

<script src="assets/js/scripts.js"></script>
