---
title: "RNA Bioinformatics exercises"
---

---------------------------------

# Introduction

All exercises for the lecture RNA Bioinformatics are provided on this web page.

The exercises complement the lecture and aim to improve your understanding of the diverse topics of this course.

:::: {#explaining .message-box }

::: {#note-exp .note-header}
```{r, include=knitr::is_html_output(), echo=FALSE,}
knitr::include_graphics("figures/infoicon.svg")
```
**Note**
:::
::: {#note-exp .note-body}


We highly recommend doing these exercises. It will help you to prepare for the
exam.

:::


::::

# Details

Exercises are meant to be done during the weekly exercise session.
During the same exercise session the results will be discussed.

Master solutions will be provided sometime after each exercise session.

:::: {#explaining .warning-box }

::: {#note-exp .warning-header}
```{r, include=knitr::is_html_output(), echo=FALSE,}
knitr::include_graphics("figures/warningicon.svg")
```
**Warning**
:::
::: {#note-exp .note-body}

You will find hints in the self learning materials which e.g. include information
about the formulae or recursion that you have to use.
However, in the exam we won't provide those formulae. You have to learn them by heart.

:::


::::


```{r, include=knitr::is_html_output(), echo=FALSE,}
knitr::asis_output('<div id="admin-panel"><a href="admin.html" class="fill-div"></a></div>')
```


